---
title: "hierarchical"
author: "Gulzina Kuttubekova"
date: "4/22/2020"
output: pdf_document
---

```{r}
library(ggplot2)
library(ggpubr)
library(dplyr)
library(reshape)
library(LearnBayes)
library(mvtnorm)
library(rmutil)
```

Read the dataset:
```{r}
covid <- read.csv("~/Desktop/spring2020/stat207/bayesian-analysis-of-COVID19-in-CA/data/Covid19-04-13-20.txt")
```

We estimate the proportions of infected people in each county:
```{r}
covid$inf_prop <- covid$Total.cases / sum(covid$Total.cases)
```

Using those proportions, find new total.cases:
```{r}
covid$new_total_cases <- round(covid$inf_prop*7.9*10^6)
covid[,c("Total.cases", "new_total_cases")]
```

Find posterior mode and hessian matrix
```{r}
inits <- c(-0.5, 0.5)
bebi_optim <- optim(par = inits, fn = betabinexch, 
                    data = covid[,c('Deaths', 'Total.cases')],
                    hessian = TRUE,
                    control = list(fnscale = -1))
```


We found posterior mode of theta1 and theta2, as well the covariance matrix:
```{r}
delta = bebi_optim$par
sigma = 2*solve(-bebi_optim$hessian)
df = 4
```

Draw samples from posterior using rejection sampling and t-distribution as a proposal distirbution with mean delta and inflated covariance matrix found from Normal approximation of log-posterior:

```{r}
# proporsal t-distribution: to find constant C
lq <- function(pars) {
    # inflate sigma
    return(dmvt(pars, delta = delta, sigma = 2*sigma, df = df))
}
```

Find log(c):
```{r}
#' Find normalizing contant = upper bound C in rejection sampling
#' 
#' @param init initial value at which lp, lq are evaluated
#' @param lp log posterior function 
#' @param lq log proporsal function
#' @return a value of log c
find_c <- function(init, lp, lq, d = covid[,c('Deaths', 'Total.cases')]) {
    
    f = function(pars) {
        return(lp(pars, data = d) - lq(pars))
    }
    
    res = optim(par = init, fn = f, control = list(fnscale = -1,
                                                  maxit = 10000000))
    
    return(res$value)
}
```

```{r}
logc <- find_c(lp = betabinexch,
            lq = lq,
            init = inits)
```


```{r}
n <- 200000
rej_sample <- rejectsampling(betabinexch, 
                             list(m = delta, var = sigma, df = df),
                             dmax = logc,
                             n = n,
                             data = covid[,c('Deaths', 'Total.cases')])

rej_est_theta <- apply(rej_sample, MARGIN = 2, FUN = mean)
rej_est_theta
```

Convert samples to previous scale:
```{r}
rej_sample[,1] <- exp(rej_sample[,1]) / (1 + exp(rej_sample[,1]))
rej_sample[,2] <- exp(rej_sample[,2])
```

```{r}
rej_ss <- data.frame(rej_sample)
colnames(rej_ss) <- c('mu', 'tau')
```

```{r}
apply(rej_ss, MARGIN = 2, FUN = mean)
```

Plot tau:
```{r}
ggdensity(rej_ss, 'tau', fill = 'blue')
```




Sample theta_i:
```{r}
sample_beta <- function(n, y, mu, tau) {
    res = rbeta(1, shape1 = mu*tau + y, shape2 = (1-mu)*tau + n - y)
    return(res)
}
```

```{r}
n = dim(rej_ss)[1]
thetas <- matrix(NA, nrow = n, ncol = 58)

for (i in 1:n) {
    for (j in 1:58)
        thetas[i,j] <- sample_beta(n = covid$Total.cases[j],
                                   y = covid$Deaths[j],
                                   mu = rej_ss$mu[i],
                                   tau = rej_ss$tau[i])
}
```

Plot thetas on the same plot:
```{r}
thetas <- data.frame(thetas)
thetas_melt <- melt(thetas)

colnames(thetas_melt) <- c('theta_i', 'value')
```


```{r}
ggdensity(thetas_melt, 'value', fill = 'theta_i') + xlim(0, 0.1) +
    rremove('legend')
```







Get estimations for each theta:
```{r}
sample_binom <- function(n, p) {
    res = rbinom(1, size = n, prob = p)
    return(res)
}
```

```{r}
y_reph <- matrix(NA, nrow = n, ncol = 58)

for (i in 1:n) {
    for (j in 1:58)
        y_reph[i,j] <- sample_binom(n = covid$new_total_cases[j],
                                    p = thetas[i,j])
}
```

```{r}
y_reph %>% apply(2, quantile, probs = c(0.025, 0.5, 0.975)) -> y_disth
y_disth <- data.frame(y_disth)
colnames(y_disth) <- covid$County
y_disth
```

```{r}
# use all 500 replicates! and calculate probabilities for each one!!! the last question
prob_count <- sum(apply(y_reph, 1, sum) >= 250000)
prob <- prob_count / n
prob
```





















