---
title: "LOO"
author: "Gulzina Kuttubekova"
date: "4/23/2020"
output: pdf_document
---

```{r}
library(ggplot2)
library(ggpubr)
library(dplyr)
library(reshape)
library(LearnBayes)
library(rmutil)
```

Read the dataset:
```{r}
covid <- read.csv("~/Desktop/spring2020/stat207/bayesian-analysis-of-COVID19-in-CA/data/Covid19-04-13-20.txt")

d <- covid[,c('Deaths', 'Total.cases')]
```

Write a function which takes out one observation at a time. We will use SIR method to sample from p(theta|y(-i)), when ith observation is removed. 

```{r}
thetas <- rej_sample
```

```{r}
i <- 1
n <- dim(thetas)[1]
w <- c()
p <- c()

# calculate w(theta_j)'s
for (j in 1:n) {
    w[j] <- betabinexch(thetas[j,], data = d[-i, ]) / betabinexch(thetas[j,], 
                                                               data = d)
}

sum_w <- sum(w)

for (j in 1:n) {
    p[j] <- w[j] / sum_w
}
```

```{r}
# sample from thetas
sample(thetas, 100, replace = TRUE, prob = p)
```
I couldn't sample thetas here, idk why.. Check later!!!!!



Do another approach: we have to sample from the p(theta|y_(-i)) and get predictive distribution of y_i: p(y_i | y_(-i)). Compare sample from posterior distribution with the observed value:
```{r}
start=c(-3,3)
fit=laplace(betabinexch,start,d[-i,])
tpar=list(m=fit$mode,var=2*fit$var,df=4)
theta=sir(betabinexch,tpar,1000,d[-i,])
```

```{r}
# sample from predictive distribution
y_1 <- c()

for (s in 1:1000) {
    y_1[s] <- rbetabinom(1, size = d[i,2], 
                         m = exp(theta[s,1]) / (1+exp(theta[s,1])),
                         s = exp(theta[s,2]))
}

```

```{r}
y_1 %>% hist()
abline(v=d[1,1], col="red")
```

```{r}
par(mfrow = c(3,4))
y_less <- c()
y_greater <- c()

for (i in 1:12) {
    start=c(-3,3)
    fit=laplace(betabinexch,start,d[-i,])
    tpar=list(m=fit$mode,var=2*fit$var,df=4)
    theta=sir(betabinexch,tpar,100,d[-i,])
    
    # sample from predictive distribution
    y_1 <- c()

    for (s in 1:100) {
    y_1[s] <- rbetabinom(1, size = d[i,2], 
                         m = exp(theta[s,1]) / (1+exp(theta[s,1])),
                         s = exp(theta[s,2]))
    }
    
    # calculate p-values that y_i is less or greater than estimated values
    y_less[i] <- sum(y_1 <= d[i,1]) / 100
    y_greater[i] <- sum(y_1 > d[i,1]) / 100
    
    # plot you findings
    y_1 %>% hist(ylab = "", main = covid$County[i])
    abline(v=d[i,1], col="red")
}
```

```{r}
# display p-values
data.frame(covid$County[1:12], y_less, y_greater)
```



```{r}
par(mfrow = c(3,4))
y_less <- c()
y_greater <- c()

for (i in 13:24) {
    start=c(-3,3)
    fit=laplace(betabinexch,start,d[-i,])
    tpar=list(m=fit$mode,var=2*fit$var,df=4)
    theta=sir(betabinexch,tpar,100,d[-i,])
    
    # sample from predictive distribution
    y_1 <- c()

    for (s in 1:100) {
    y_1[s] <- rbetabinom(1, size = d[i,2], 
                         m = exp(theta[s,1]) / (1+exp(theta[s,1])),
                         s = exp(theta[s,2]))
    }
    
    # calculate p-values that y_i is less or greater than estimated values
    y_less[i] <- sum(y_1 <= d[i,1]) / 100
    y_greater[i] <- sum(y_1 > d[i,1]) / 100
    
    # plot
    y_1 %>% hist(ylab = "", main = covid$County[i])
    abline(v=d[i,1], col="red")
}
```

```{r}
# display p-values
data.frame(covid$County[13:24], y_less[13:24], y_greater[13:24])
```


```{r}
par(mfrow = c(3,4))
y_less <- c()
y_greater <- c()

for (i in 25:36) {
    start=c(-3,3)
    fit=laplace(betabinexch,start,d[-i,])
    tpar=list(m=fit$mode,var=2*fit$var,df=4)
    theta=sir(betabinexch,tpar,100,d[-i,])
    
    # sample from predictive distribution
    y_1 <- c()

    for (s in 1:100) {
    y_1[s] <- rbetabinom(1, size = d[i,2], 
                         m = exp(theta[s,1]) / (1+exp(theta[s,1])),
                         s = exp(theta[s,2]))
    }
    
    # calculate p-values that y_i is less or greater than estimated values
    y_less[i] <- sum(y_1 <= d[i,1]) / 100
    y_greater[i] <- sum(y_1 > d[i,1]) / 100
    
    # plot
    y_1 %>% hist(ylab = "", main = covid$County[i])
    abline(v=d[i,1], col="red")
}
```

```{r}
# display p-values
data.frame(covid$County[25:36], y_less[25:36], y_greater[25:36])
```



```{r}
par(mfrow = c(3,4))
y_less <- c()
y_greater <- c()

for (i in 37:48) {
    start=c(-3,3)
    fit=laplace(betabinexch,start,d[-i,])
    tpar=list(m=fit$mode,var=2*fit$var,df=4)
    theta=sir(betabinexch,tpar,100,d[-i,])
    
    # sample from predictive distribution
    y_1 <- c()

    for (s in 1:100) {
    y_1[s] <- rbetabinom(1, size = d[i,2], 
                         m = exp(theta[s,1]) / (1+exp(theta[s,1])),
                         s = exp(theta[s,2]))
    }
    
    # calculate p-values that y_i is less or greater than estimated values
    y_less[i] <- sum(y_1 <= d[i,1]) / 100
    y_greater[i] <- sum(y_1 > d[i,1]) / 100
    
    # plot
    y_1 %>% hist(ylab = "", main = covid$County[i])
    abline(v=d[i,1], col="red")
}
```

```{r}
# display p-values
data.frame(covid$County[37:48], y_less[37:48], y_greater[37:48])
```


```{r}
par(mfrow = c(3,4))
y_less <- c()
y_greater <- c()

for (i in 49:58) {
    start=c(-3,3)
    fit=laplace(betabinexch,start,d[-i,])
    tpar=list(m=fit$mode,var=2*fit$var,df=4)
    theta=sir(betabinexch,tpar,100,d[-i,])
    
    # sample from predictive distribution
    y_1 <- c()

    for (s in 1:100) {
    y_1[s] <- rbetabinom(1, size = d[i,2], 
                         m = exp(theta[s,1]) / (1+exp(theta[s,1])),
                         s = exp(theta[s,2]))
    }
    
    # calculate p-values that y_i is less or greater than estimated values
    y_less[i] <- sum(y_1 <= d[i,1]) / 100
    y_greater[i] <- sum(y_1 > d[i,1]) / 100
    
    # plot
    y_1 %>% hist(ylab = "", main = covid$County[i])
    abline(v=d[i,1], col="red")
}
```

```{r}
# display p-values
data.frame(covid$County[49:58], y_less[49:58], y_greater[49:58])
```

Pick only the ones with the highest deviations:
```{r}
par(mfrow = c(3,3))

for (i in c(12, 14:16, 18, 30, 34, 35, 37)) {
    start=c(-3,3)
    fit=laplace(betabinexch,start,d[-i,])
    tpar=list(m=fit$mode,var=2*fit$var,df=4)
    theta=sir(betabinexch,tpar,2000,d[-i,])
    
    # sample from predictive distribution
    y_1 <- c()

    for (s in 1:2000) {
    y_1[s] <- rbetabinom(1, size = d[i,2], 
                         m = exp(theta[s,1]) / (1+exp(theta[s,1])),
                         s = exp(theta[s,2]))
    }
    
    # calculate p-values that y_i is less or greater than estimated values
    y_less[i] <- sum(y_1 <= d[i,1]) / 2000
    y_greater[i] <- sum(y_1 > d[i,1]) / 2000
    
    # plot
    y_1 %>% hist(ylab = "", main = covid$County[i], xlab = 'death toll')
    abline(v=d[i,1], col="red")
}
```















